A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2019-02-01 paper/submission/ZW10157-response-1.tex
lot's of small edits, mostly commas

{\bf Our response:} {\color{blue} A line XX has been added indicating this capability in GPT too.}

@@

{\bf Our response:} {\color{blue} 
	Respectfully, we point out Referee A's comment that 
	``Computing and algorithms'' is listed under PRAB's scope. 
	However, for those interested in physics and not implementation details (i.e. user of code not developer), 
	we had hoped the AWA optimization problem would 
	provide a meaningful and tangible use case for the framework.
	It seems we did not present that part of the paper strongly enough. 
	We hope some of the additions above have helped in this area.
	However, would also like to keep some computational details for those who 
	may be interested in code development specifically for accelerator optimization.
}

@@

{\bf Our response: }{\color{blue} 	
	We hope the additions above are addressing these issues. }

>>>>

{\bf Our response:} {\color{blue} Line 123 in the new manuscript indicates this capability is in GPT too.}

@@

{\bf Our response:} {\color{blue} 
	Additions were made to the text to enhance the discussion of the 
	physical problem and extend the discussion of the results.  For example, 
	the abstract was extended, and additional discussion was added to the 
	optimization results section. The code snippets were moved out of the 
	main body of the paper.  However, we would also like to keep some 
	computational details for those who may be interested in code 
	development specifically for accelerator optimization.  Hopefully this 
	is justified since "Computing and algorithms" is listed under PRAB's 
	scope.
}

@@

{\bf Our response: }{\color{blue} 	
	We hope the additions above (including response to Ref. A) are addressing these issues.}

### 2019-02-01 paper/submission/prab.tex
new bib, small changes to paper
lot's of small edits, mostly commas

\begin{abstract}
Particle accelerators are invaluable tools for research in the basic and applied sciences, such as materials science, chemistry,
  the biosciences, particle physics, nuclear physics and medicine. The design, commissioning, and operation of accelerator facilities is a
  non-trivial task, due to the large number of control parameters and the complex interplay of several conflicting design goals.
  The Argonne Wakefield Accelerator facility has some unique challenges resulting from its purpose to carry out advanced accelerator R\&D.
  Individual experiments often have challenging beam requirements, and the physical configuration of the beamlines is often changed
  to accommodate the variety of supported experiments. The need for rapid deployment of different operational settings
  further complicates the optimization work that must be done for multiple constraints and challenging operational regimes. 
  One example of this is an independently staged two-beam acceleration experiment which requires the construction 
  of an additional beamline (this is now in progress).  The high charge drive beam, well into the space charge regime, must be threaded
  through small aperture (\SI{17.6}{mm}) decelerating structures.  
  In addition, the bunch length must be sufficiently short to maximize power generation in the decelerator.  
We propose to tackle this problem by means of multi-objective optimization algorithms which also facilitate a parallel deployment.
In order to compute solutions in a meaningful time frame, a fast and scalable software framework is required.
In this paper, we present a general-purpose framework for simulation-based
multi-objective optimization methods that allows the automatic investigation of optimal sets of machine parameters.
Using evolutionary algorithms as the optimizer and \textsc{OPAL} as the forward solver, validation experiments and 
 results of multi-objective optimization problems in the domain of beam dynamics are presented. 
  Optimized solutions for the new high charge drive beamline found by the framework were used to finish the design 
  of a two beam acceleration experiment.
  The selected solution along with the associated beam parameters is presented.

@@

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research and practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.
Despite the success of on-line models in some facilities~\cite{xiaobiao}, 
and various model dependent and model independent tuning and optimization techniques, 
empirical tuning by operators is a common method used at many research facilities.
When the beam dynamics is non-linear, as is the case with space charge, 
simple and fast models are applicable only in a very restricted manner. 
When it comes to multi-objective optimization, the situation is even more complicated.
In order to be able to reliably identify optimal configurations of
  accelerators, we solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning parameters.
  This approach has been used in the past with much success~\cite{hofler13,bazarov05,yrss:09,jefferson,gull1,gull2}.
The difference here being the implementation and application to a 
problem at the Argonne Wakefield Accelerator Facility (AWA).

A hallmark of the AWA facility is the flexibility to swap physical
components in the beamlines, which enables different, often novel,
accelerator research experiments to take place.  Not only do the
physical machine components change, the beam characteristics also vary
considerably to meet different needs.  The facility operates at both low
and high charge (up to 100 nC), and at high charge strong nonlinearities
require a full 3D space charge approach in simulations for modeling.  
Finding optics solutions in this regime, especially when
there are additional constraints such as the small aperture two-beam
accelerating structures, is challenging even without the quick
turnaround of the beamline configurations.  Therefore, it has been an
important research objective to develop a precise, e.g. 3D model
embedded into a multi-objective optimization framework that may be used
as a flexible platform for optimization of changing machine
configurations operated at different charge levels. While other codes,
such as GPT~\cite{gpt} and ELEGANT~\cite{elegant}, also have
integrated genetic optimization algorithms; the OPAL~\cite{opal} framework
differentiates itself by being open source (i.e. free to use), massively
parallel, and fully 3D.

@@

The framework discussed here, incorporates the following three contributions:
%
\begin{enumerate}
  \item Implementation of a scalable optimization algorithm capable of
        approximating Pareto fronts in high dimensional spaces,
  \item design and implementation of a modular framework that is simple to use
        and deploy on large scale computational resources, and
  \item demonstration of the usefulness of the proposed framework on a real world
        application in the domain of particle accelerators. This is done
        with the optimization problem set as 
        the high charge photoinjector at the Argonne Wakefield Accelerator (AWA). 
\end{enumerate}

@@

Optimization problems deal with finding one or more feasible solutions
  corresponding to extreme values of objectives.
If more than one objective is present in the optimization problem this is called
  a multi-objective optimization problem (MOOP).
A MOOP is defined as
%
\begin{align}
  \text{ min} \quad & \quad f_m({\bf x}), ~& m &= 1, \dots, M, \label{eq:moop:obj}\\
  \text{s.t.} \quad & \quad g_j({\bf x}) \geq 0, & j &= 0, \dots, J,
  \label{eq:moop:constr}\\
  \quad & \quad  x_i^L \leq {\bf x}=x_i \leq x_i^U,& i &=0, \dots, n
  \label{eq:moop:dvar} \text{,}
\end{align}
%
where $\bf f$ denotes the objectives (\ref{eq:moop:obj}),
  $\bf g$ the constraints (\ref{eq:moop:constr}),
  and $\bf x$ the design variables (\ref{eq:moop:dvar}).
Often, conflicting objectives are encountered, and this complicates the concept of
  optimality. \textit{Pareto optimality} is often used in such situations [].
The set of Pareto optimal points forms the Pareto front or
  surface.
All points on this surface are considered to be Pareto optimal.
Sampling Pareto fronts is far from trivial.
A number of approaches have been proposed,
  e.g.\ evolutionary algorithms~\cite{deb:09},
  simulated annealing~\cite{kigv:83},
  swarm methods~\cite{keeb:95},
  and many more~\cite{domc:96,cati:02,kara:05,hoss:09}.
In the next section, we briefly introduce the theory of evolutionary algorithms
  used in the present work.

@@

Applied to multi-objective optimization problems, genes correspond to
  design variables.
The fitness of an individual is loosely related to the value of the objective
  functions for the corresponding genes.
Figure~\ref{fig:varsel} schematically depicts the connection of the
  components introduced above.
The process starts with an initially random population of individuals, each
  individual with a unique set of genes and corresponding fitness,
  representing one location in the search space.
In a next step the population is processed by the selector
  determining the $k$ fittest individuals according to their fitness values.
While the $k$ fittest individuals are passed to the variator, the
  remaining $n-k$ individuals are eliminated from the population.
The \textsc{Variator} mates and recombines the $k$ fittest individuals to
  generate new offspring.
After evaluating the fitness of all the freshly born individuals a
  \textit{generation} cycle has completed and the process can start anew.

@@

Simulation based multi-objective optimization problems are omnipresent in
  research and industry.
The simulation and optimization problems arising in such problems are in
  general very big and computationally demanding.
This motivated us to design a massively parallel general purpose framework.
The key traits of such a design can be summarized as:
%
\begin{itemize}
  \item support any multi-objective optimization method,
  \item support any function evaluator: simulation code or measurements,
  \item offer a general description/specification of objectives, constraints
        and design variables,
  \item run efficiently in parallel on current large-scale high-end clusters
        and supercomputers.
\end{itemize}
%

@@

Several similar frameworks, e.g.~\cite{fide:09,lems:09,lbjt:07,dnld:06}, have
  been proposed.
Commonly these frameworks are tightly coupled to an optimization algorithm,
  e.g.\ only providing evolutionary algorithms as optimizers.
Users can merely specify optimization problems, but cannot change the
  optimization algorithm.
Our framework follows a more general approach, providing a user-friendly way
  to introduce new or choose from existing built-in multi-objective
  optimization algorithms.
Tailoring the optimization algorithm to the optimization problem at hand is
  an important feature due to the many different characteristics of
  optimization problems that should be handled by such a general framework.
As an example, it is shown how \textsc{Pisa} \cite{pisa}, an existing evolutionary
  algorithm library, was integrated with ease.
Similarly, other multi-objective algorithms could be incorporated and
  used to solve optimization problems.

@@

Every available MPI process will take up one of the three available roles (see
  Fig.~\ref{fig:framenetwork}):  one process acts as \textsc{Pilot}, the
  remaining processes are divided amongst \textsc{Worker} and
  \textsc{Optimizer} roles.
Both, the \textsc{Worker} and the \textsc{Optimizer} can consist of multiple
  MPI processes to exploit parallelism.
As shown in Fig.~\ref{fig:opt-framework-layout} the \textsc{Pilot} is used
  to coordinate all ``information requests'' between the \textsc{Optimizer}
  and the \textsc{Worker}.
An information request is a job that consists of a set of design variables
  (e.g.~the genes of an individual) and a type of information it requests
  (e.g.~function evaluation or derivative).
The \textsc{Pilot} keeps checking for idle \textsc{Worker} and assigns jobs
  in the queue to any free \textsc{Worker}.
Once the \textsc{Worker} has computed and evaluated the request its results
  are returned to the \textsc{Optimizer} that originally requested the
  information.

After a process gets appointed a role it starts a polling loop asynchronously
  checking for appropriate incoming requests.
To that end a \textsc{Poller} interface helper class has been introduced.
The \textsc{Poller} interface consists of an infinite loop that checks
  periodically for new MPI messages.
Upon reception a new message is immediately forwarded to the appropriate
  handler: the \texttt{onMessage()} method.
The method is called with the \texttt{MPI\_Status} of the received message and
  a \texttt{size\_t} value specifying different values depending on the value
  of the \texttt{MPI\_Tag}.
The \textsc{Poller} interface allows the implementation of special methods
  (denoted \textit{hooks}) determining the behavior of the polling process,
  e.g.\ for actions that need to be taken after a message has been handled.
Every \textsc{Poller} terminates the loop upon receiving a special MPI tag.

@@

All processors running an \textsc{Optimizer} component call the
  \texttt{initialize} entry point after role assignment in the
  \textsc{Pilot}.
The implementation of \texttt{initialize} must set up and start the poller and
  the optimization code.
Since an optimizer derives from the \texttt{Poller} interface, predefined
  hooks can be used to determine the polling procedure.
Hooks can be implemented as empty methods, but the \texttt{onMessage}
  implementation should reflect the optimization part of the protocol for
  handling events from the \textsc{Pilot}.
A special set of communicator groups serves as communication channels to the
  \textsc{Pilot} and its job queue and if existing to processes supporting the
  \textsc{Optimizer} component.

@@

In most cases forward solver implementations are simple wrappers to run
  an existing ``external'' simulation code using a set of design variables as
  input. In case of the OPAL integration, basically the \texttt{main} function is
  playing the role of the ``forward solver''. To underline the general nature of our approach, 
  in a similar project, the described methods are used for cavity shape optimisation based on \cite{ARBENZ2008381}. 
As for the \textsc{Optimizer} component there exists a base class, labeled
  \texttt{Simulation} as common basis for all \textsc{Simulation}
  implementations.
In addition, this component also inherits from the \texttt{Worker} class,
  already implementing the polling protocol for default worker types.
As shown in the API in Listing 3, the \texttt{Worker} class expects an
  implementation to provide implementations for those three methods.

@@

We aimed at an easy and expressive way for users to specify multi-objective
  optimization problems.
Following the principle of keeping metadata (optimization and simulation
  input data) together, we decided to embed the optimization problem
  specification in the simulation input file by prefixing it with special
  characters, e.g. as annotations prefixed with a special character.
In some cases it might not be possible to annotate the simulation input file.
By providing an extra input file parser, optimization problems can be read
  from stand-alone files.

To allow arbitrary constraints and objective expressions, such as
%
%{0.2cm}
\begin{Verbatim}[fontsize=\scriptsize]
  name: OBJECTIVE,
        EXPR="5 * average(42.0, "measurement.dat") + ENERGY";
\end{Verbatim}
%\vspace{0.2cm}
%
\noindent
  An expression parser using Boost Spirit~\cite{boost} was implemented.
In addition to the parser, we need an evaluator able to evaluate an expression,
  given a parse tree and variable assignments to an actual value.
Expressions arising in multi-objective optimization problems usually evaluate
  to booleans or floating point values.
The parse tree, also denoted abstract syntax tree (AST), is constructed
  recursively while an expression is parsed.
Upon evaluation, all unknown variables are replaced with values, either
  obtained from simulation results or provided by other subtrees in the AST.
In this stage, the AST can be evaluated bottom-up and the desired result is
  returned after processing the root of the tree.

To improve the expressive power of objectives and constraints, a
  simple mechanism to define and call custom functions in expressions was introduced.
Using simple functors, e.g. as the one shown in Listing 4 to compute an
  average over a set of data points, enriches expressions with custom
  functions.
Custom function implementations overload the \texttt{()} parenthesis operator.
The function arguments specified in the corresponding expression are stored in
  a \texttt{std::vector} of Boost variants~\cite{boost2} that can be
  booleans, strings or floating point values.

@@

One major disadvantage of the master/slave implementation model is the fast
  saturation of the network links surrounding the master node.
In \cite{bctg:09} authors observe an exponential increase in hot-spot latency
  with increasing number of workers that are attached to one master process.
Clearly, the limiting factor is the number of outgoing links of a node in the
  network topology and already for a few workers the links surrounding a
  master process are subject to congestions.
This effect is amplified further by large message sizes.

To that end we implemented a solution propagation based on rumor networks 
(see \cite{bgps:06,ayss:09}) using only one-sided communication.
This limits the number of messages sent over the already heavily used links
  surrounding the master node and at the same time helps to prevent the use of
  global communication.
Using information about the interconnection network topology and the
  application communication graph the task of assigning roles helps to further
  improve the parallel performance.

@@

The framework contains a wrapper implementing the API mentioned in
  Listing 3 for using \textsc{OPAL} \cite{opal} as the forward solver.
\textsc{OPAL} provides different trackers for cyclotrons and linear
  accelerators with satisfactory parallel performance.

With access to the \textsc{OPAL} forward solver the framework is able to
  tackle a multitude of optimization problems arising in the domain of
  particle accelerators.
  The framework is also integrated into \textsc{OPAL} so that users can 
  define optimization problems within an input file, requiring no 
  additional knowledge or installation of the API to use it.

@@

If the objectives and constraints are simple arithmetical expressions, the
  \texttt{FunctionEvaluator} simulator can be used.
Using functors and the default expression primitives already powerful
  multi-objective optimization problems can be specified, i.e.\ the benchmark
  problem presented in \cite{hbwh:05}:
%

%\begin{strip}
	\begin{widetext}
		\begin{align} \label{eqn:bench}
		\text{min} & \left[ 1 - \exp \left( -1 \left(
		\left(x_1 - \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_2 - \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_3 - \frac{1}{\sqrt{3}} \right)^2 \right)\right), \right. \\
		\vspace{3em} 
		& \left. 1 - \exp \left( -1 \left(
		\left(x_1 + \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_2 + \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_3 + \frac{1}{\sqrt{3}} \right)^2 \right)\right) \right]^T \nonumber \\
		\vspace{3em} \nonumber \\
		\text{s.t.} \quad & \quad -1 \le x_i \le 1, \quad i=1,2,3 \nonumber
		\text{.} \nonumber
		\end{align}
	\end{widetext}
%\end{strip}

>>>>

\begin{abstract}
Particle accelerators are invaluable tools for research in the basic and applied sciences, such as materials science, chemistry,
  the biosciences, particle physics, nuclear physics and medicine. The design, commissioning, and operation of accelerator facilities is a
  non-trivial task, due to the large number of control parameters and the complex interplay of several conflicting design goals.
  The Argonne Wakefield Accelerator facility has some unique challenges resulting from its purpose to carry out advanced accelerator R\&D.
  Individual experiments often have challenging beam requirements, and the physical configuration of the beamlines is often changed
  to accommodate the variety of supported experiments. The need for rapid deployment of different operational settings
  further complicates the optimization work that must be done for multiple constraints and challenging operational regimes. 
  One example of this is an independently staged two-beam acceleration experiment which requires the construction 
  of an additional beamline (this is now in progress).  The high charge drive beam, well into the space charge regime, must be threaded
  through small aperture (17.6 mm) decelerating structures.  
  In addition, the bunch length must be sufficiently short to maximize power generation in the decelerator.  
We propose to tackle this problem by means of multi-objective optimization algorithms which also facilitate a parallel deployment.
In order to compute solutions in a meaningful time frame, a fast and scalable software framework is required.
In this paper, we present a general-purpose framework for simulation-based
multi-objective optimization methods that allows the automatic investigation of optimal sets of machine parameters.
Using evolutionary algorithms as the optimizer and \textsc{OPAL} as the forward solver, validation experiments and 
 results of multi-objective optimization problems in the domain of beam dynamics are presented. 
  Optimized solutions for the new high charge drive beamline found by the framework were used to finish the design 
  of a two beam acceleration experiment.
  The selected solution along with the associated beam parameters is presented.

@@

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research and practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.
Despite the success of on-line models in some facilities~\cite{xiaobiao}, 
and various model dependent and model independent tuning and optimization techniques, 
empirical tuning by operators is a common method used at many research facilities.
When the beam dynamics is nonlinear, as is the case with space charge, 
simple and fast models are applicable only in a very restricted manner. 
This further complicates any mult-objective optimization by complicating the model. 
In order to be able to reliably identify optimal configurations of
  accelerators, we solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning parameters.
  This approach has been used in the past with much success~\cite{hofler13,bazarov05,yrss:09,jefferson,gull1,gull2,marija}.
The difference here being the implementation and application to a 
problem at the Argonne Wakefield Accelerator Facility (AWA).

A hallmark of the AWA facility is the flexibility to swap physical
components in the beamlines, which enables different, often novel,
accelerator research experiments to take place.  Not only do the
physical machine components change, the beam characteristics also vary
considerably to meet different needs.  The facility operates at both low
and high charge (up to 100 nC), and at high charge strong nonlinearities
require a full 3D space charge approach in simulations.  
Finding optics solutions in this regime, especially when
there are additional constraints such as the small aperture two-beam
accelerating structures, is challenging even without the quick
turnaround of the beamline configurations.  Therefore, it has been an
important research objective to develop a precise, e.g. 3D model
embedded into a multi-objective optimization framework that may be used
as a flexible platform for optimization of changing machine
configurations operated at different charge levels. While other codes,
such as GPT~\cite{gpt} and ELEGANT~\cite{elegant}, also have
integrated genetic optimization algorithms; the OPAL~\cite{opal} framework
differentiates itself by being open source (i.e. free to use), massively
parallel, and fully 3D.

@@

The framework discussed here, incorporates the following three contributions:
%
\begin{enumerate}
  \item Implementation of a scalable optimization algorithm capable of
        approximating Pareto fronts in high dimensional spaces,
  \item design and implementation of a modular framework that is simple to use
        and deploy on large scale computational resources, and
  \item demonstration of the usefulness of the proposed framework on a real world
        application in the domain of particle accelerators. This is done
        with the optimization problem set as 
        the high charge photoinjector at the AWA. 
\end{enumerate}

@@

Optimization problems deal with finding one or more feasible solutions
  corresponding to extreme values of objectives.
If more than one objective is present in the optimization problem this is called
  a multi-objective optimization problem (MOOP).
A MOOP is defined as
%
\begin{align}
  \text{ min} \quad & \quad f_m({\bf x}), ~& m &= 1, \dots, M, \label{eq:moop:obj}\\
  \text{s.t.} \quad & \quad g_j({\bf x}) \geq 0, & j &= 0, \dots, J,
  \label{eq:moop:constr}\\
  \quad & \quad  x_i^L \leq {\bf x}=x_i \leq x_i^U,& i &=0, \dots, n
  \label{eq:moop:dvar} \text{,}
\end{align}
%
where $\bf f$ denotes the objectives (\ref{eq:moop:obj}),
  $\bf g$ the constraints (\ref{eq:moop:constr}),
  and $\bf x$ the design variables (\ref{eq:moop:dvar}).
Often, conflicting objectives are encountered, and this complicates the concept of
  optimality. \textit{Pareto optimality} is often used in such situations.
The set of Pareto optimal points forms the Pareto front or
  surface.
All points on this surface are considered to be Pareto optimal.
Sampling Pareto fronts is far from trivial.
A number of approaches have been proposed,
  e.g.\ evolutionary algorithms~\cite{deb:09},
  simulated annealing~\cite{kigv:83},
  swarm methods~\cite{keeb:95},
  and many more~\cite{domc:96,cati:02,kara:05,hoss:09}.
In the next section, we briefly introduce the theory of evolutionary algorithms
  used in the present work.

@@

Applied to multi-objective optimization problems, genes correspond to
  design variables.
The fitness of an individual is loosely related to the value of the objective
  functions for the corresponding genes.
Figure~\ref{fig:varsel} schematically depicts the connection of the
  components introduced above.
The process starts with an initially random population of individuals, each
  individual with a unique set of genes and corresponding fitness,
  representing one location in the search space.
In the next step, the population is processed by the selector
  determining the $k$ fittest individuals according to their fitness values.
While the $k$ fittest individuals are passed to the variator, the
  remaining $n-k$ individuals are eliminated from the population.
The \textsc{Variator} mates and recombines the $k$ fittest individuals to
  generate new offspring.
After evaluating the fitness of all the freshly born individuals a
  \textit{generation} cycle has completed and the process can start anew.

@@

Simulation based multi-objective optimization problems are omnipresent in
  research and industry.
These simulation and optimization problems are in
  general very big and computationally demanding.
This motivated us to design a massively parallel general purpose framework.
The key traits of such a design can be summarized as:
%
\begin{itemize}
  \item support any multi-objective optimization method,
  \item support any function evaluator: simulation code or measurements,
  \item offer a general description/specification of objectives, constraints
        and design variables,
  \item run efficiently in parallel on current large-scale high-end clusters
        and supercomputers.
\end{itemize}
%

@@

Several similar frameworks, e.g.~\cite{fide:09,lems:09,lbjt:07,dnld:06}, have
  been proposed.
Commonly these frameworks are tightly coupled to an optimization algorithm,
  e.g.\ only providing evolutionary algorithms as optimizers.
Users can specify optimization problems, but cannot change the
  optimization algorithm.
Our framework follows a more general approach, providing a user-friendly way
  to introduce new or choose from existing built-in multi-objective
  optimization algorithms.
Tailoring the optimization algorithm to the optimization problem at hand is
  an important feature due to the many different characteristics of
  optimization problems that should be handled by such a general framework.
As an example, it is shown how \textsc{Pisa}~\cite{pisa}, an existing evolutionary
  algorithm library, was integrated with ease.
Similarly, other multi-objective algorithms could be incorporated and
  used to solve optimization problems.

@@

Every available MPI process will take up one of the three available roles (see
  Fig.~\ref{fig:framenetwork}):  one process acts as \textsc{Pilot}, the
  remaining processes are divided amongst \textsc{Worker} and
  \textsc{Optimizer} roles.
Both, the \textsc{Worker} and the \textsc{Optimizer} can consist of multiple
  MPI processes to exploit parallelism.
As shown in Fig.~\ref{fig:opt-framework-layout}, the \textsc{Pilot} is used
  to coordinate all ``information requests'' between the \textsc{Optimizer}
  and the \textsc{Worker}.
An information request is a job that consists of a set of design variables
  (e.g.~the genes of an individual) and a type of information it requests
  (e.g.~function evaluation or derivative).
The \textsc{Pilot} keeps checking for idle \textsc{Worker} and assigns jobs
  in the queue to any free \textsc{Worker}.
Once the \textsc{Worker} has computed and evaluated the request its results
  are returned to the \textsc{Optimizer} that originally requested the
  information.

After a process gets appointed a role, it starts a polling loop to asynchronously
  check for appropriate incoming requests.
To that end a \textsc{Poller} interface helper class has been introduced.
The \textsc{Poller} interface consists of an infinite loop that checks
  periodically for new MPI messages.
Upon reception a new message is immediately forwarded to the appropriate
  handler: the \texttt{onMessage()} method.
The method is called with the \texttt{MPI\_Status} of the received message and
  a \texttt{size\_t} value specifying different values depending on the value
  of the \texttt{MPI\_Tag}.
The \textsc{Poller} interface allows the implementation of special methods
  (denoted \textit{hooks}) determining the behavior of the polling process,
  e.g.\ for actions that need to be taken after a message has been handled.
Every \textsc{Poller} terminates the loop upon receiving a special MPI tag.

@@

All processors running an \textsc{Optimizer} component call the
  \texttt{initialize} entry point after role assignment in the
  \textsc{Pilot}.
The implementation of \texttt{initialize} must set up and start the poller and
  the optimization code.
Since an optimizer derives from the \texttt{Poller} interface, predefined
  hooks can be used to determine the polling procedure.
Hooks can be implemented as empty methods, but the \texttt{onMessage}
  implementation should reflect the optimization part of the protocol for
  handling events from the \textsc{Pilot}.
A special set of communicator groups serves as communication channels to the
  \textsc{Pilot}, its job queue, and processes supporting the
  \textsc{Optimizer} component.

@@

In most cases, forward solver implementations are simple wrappers to run
  an existing ``external'' simulation code using a set of design variables as
  input. In the case of the OPAL integration, the \texttt{main} function is
  playing the role of the ``forward solver''. To underline the general nature of our approach, 
  in a similar project, the described methods are used for cavity shape optimisation based on \cite{ARBENZ2008381}. 
As for the \textsc{Optimizer} component there exists a base class, labeled
  \texttt{Simulation} as common basis for all \textsc{Simulation}
  implementations.
In addition, this component also inherits from the \texttt{Worker} class,
  already implementing the polling protocol for default worker types.
As shown in the API in Listing 3, the \texttt{Worker} class expects an
  implementation to provide implementations for those three methods.

@@

We aimed at an easy and expressive way for users to specify multi-objective optimization problems.
Following the principle of keeping metadata (optimization and simulation input data) together, 
we decided to embed the optimization problem specification in the simulation input file by 
prefixing it with special characters, e.g. as annotations prefixed with a special character.
In some cases, it might not be possible to annotate the simulation input file.
By providing an extra input file parser, optimization problems can be read from stand-alone files.
To allow arbitrary constraints and objective expressions, such as
%
%{0.2cm}
\begin{Verbatim}[fontsize=\scriptsize]
  name: OBJECTIVE,
        EXPR="5 * average(42.0, "measurement.dat") + ENERGY";
\end{Verbatim}
%\vspace{0.2cm}
%
\noindent
An expression parser using Boost Spirit~\cite{boost} was implemented.
In addition to the parser, we need an evaluator able to evaluate an expression,
given a parse tree and variable assignments to an actual value.
Expressions arising in multi-objective optimization problems usually evaluate
to booleans or floating point values.
The parse tree, also denoted abstract syntax tree (AST), is constructed recursively while an expression is parsed.
Upon evaluation, all unknown variables are replaced with values, 
either obtained from simulation results or provided by other subtrees in the AST.
In this stage, the AST can be evaluated bottom-up and the desired result is
  returned after processing the root of the tree.

To improve the expressive power of objectives and constraints, a
  simple mechanism to define and call custom functions in expressions was introduced.
Using simple functors, e.g. as the one shown in Listing 4, to compute an
  average over a set of data points, enriches expressions with custom
  functions.
Custom function implementations overload the \texttt{()} parenthesis operator.
The function arguments specified in the corresponding expression are stored in
  a \texttt{std::vector} of Boost variants~\cite{boost2} that can be
  booleans, strings or floating point values.

@@

One major disadvantage of the master/slave implementation model is the fast
  saturation of the network links surrounding the master node.
In \cite{bctg:09} authors observe an exponential increase in hot-spot latency
  with increasing number of workers that are attached to one master process.
The limiting factor is the number of outgoing links of a node in the
  network topology. For a few workers, the links surrounding a
  master process are subject to congestion.
This effect is amplified further by large message sizes.

To that end we implemented a solution propagation based on rumor networks 
(see \cite{bgps:06,ayss:09}) using only one-sided communication.
This limits the number of messages sent over the already heavily used links
surrounding the master node and helps to prevent the use of
global communication. Using information about the interconnection network topology and the
application communication graph, the task of assigning roles helps to further
improve the parallel performance.

@@

The framework contains a wrapper implementing the API mentioned in
  Listing 3 for using \textsc{OPAL}~\cite{opal} as the forward solver.
\textsc{OPAL} provides different trackers for cyclotrons and linear
  accelerators with satisfactory parallel performance. 
With access to the \textsc{OPAL} forward solver, the framework is able to
  tackle a multitude of optimization problems arising in the domain of
  particle accelerators.
  The framework is also integrated into \textsc{OPAL} so that users can 
  define optimization problems within an input file, requiring no 
  additional knowledge or installation of the API to use it.

@@

If the objectives and constraints are simple arithmetical expressions, 
the \texttt{FunctionEvaluator} simulator can be used.
Using functors and the default expression primitives, 
multi-objective optimization problems can be specified, 
i.e.\ the benchmark problem presented in \cite{hbwh:05}:
%

	\begin{widetext}
		\begin{align} \label{eqn:bench}
		\text{min} & \left[ 1 - \exp \left( -1 \left(
		\left(x_1 - \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_2 - \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_3 - \frac{1}{\sqrt{3}} \right)^2 \right)\right), \right. \\
		\vspace{3em} 
		& \left. 1 - \exp \left( -1 \left(
		\left(x_1 + \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_2 + \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_3 + \frac{1}{\sqrt{3}} \right)^2 \right)\right) \right]^T \nonumber \\
		\vspace{3em} \nonumber \\
		\text{s.t.} \quad & \quad -1 \le x_i \le 1, \quad i=1,2,3 \nonumber
		\text{.} \nonumber
		\end{align}
	\end{widetext}
%\end{strip}
